Arhangelskii's \(\alpha\)-principles and selection games

### 2019-11-06 articles/selection-alpha-games/article.tex
discuss alpha_1.5

\begin{proposition}
Let \(\mc A\) be \(\Gamma\)-like.
For \(i\in\{2,3,4\}\), \(\alpha_i(\mc A,\mc B)\) is equivalent to
\(\alpha_{i.1}(\mc A,\mc B)\).
\end{proposition}

@@

\begin{proposition}
Let \(\mc A\) be a \(\Gamma\)-like collection closed under finite unions.
Then \(\alpha_1(\mc A,\mc A)\) is equivalent to
\(\alpha_{1.1}(\mc A,\mc A)\).
\end{proposition}

\begin{proof}
Let \(A_n\in\mc A\) and assume \(\alpha_{1.1}(\mc A,\mc A)\).
To apply the assumption, we will define a pairwise disjoint
collection \(\{A_n':n<\omega\}\). First let \(0'=0\) and \(A_0'=A_0\).
Then suppose \(m'\geq m\) and \(A_m'\subseteq A_{m'}\subseteq\bigcup_{i\leq m}A_i'\) 
are defined for all \(m\leq n\).

If \(A_k\setminus\bigcup_{m\leq n}A_m'\) is finite for \(k>n'\), let
\(B=\bigcup_{m\leq n'}A_m\in\mc A\). This \(B\) then witnesses 
\(\alpha_1(\mc A,\mc A)\) since \(A_k\setminus B\) is finite for all \(k<\omega\).

Otherwise pick the minimal \((n+1)'>n\) where 
\(A_{n+1}'=A_{(n+1)'}\setminus\bigcup_{m\leq n}A_m'\) is infinite.
It follows that \(A_{n+1}'\subseteq A_{(n+1)'}\subseteq \bigcup_{m\leq n+1}A_m'\).
By construction, \(\{A_n':n<\omega\}\) is a pairwise disjoint collection of
members of \(\mc A\), and we may
apply \(\alpha_{1.1}(\mc A,\mc A)\) to obtain \(B\in\mc A\) where
\(A_n'\setminus B\) is finite for all \(n<\omega\).

@@

TODO add discussion of \(\alpha_{1.5}\) here.

@@

\begin{question}
Let \(\mc A,\mc B\) be strongly-\(\Gamma\)-like
and \(\mc B\subseteq\mc A\).
Is it consistent that \(\alpha_2(\mc A,\mc B)\)
implies \(\alpha_1(\mc A,\mc B)\)?
\end{question}

@@

The author would like to thank Dr. Alan Dow and Dr. Jared Holshouser
for their discussions on the subject of the paper.

>>>>

\begin{proposition}\label{pointOne}
Let \(\mc A\) be \(\Gamma\)-like.
For \(i\in\{2,3,4\}\), \(\alpha_i(\mc A,\mc B)\) is equivalent to
\(\alpha_{i.1}(\mc A,\mc B)\).
\end{proposition}

@@

\begin{theorem}
Let \(\mc A\) be a \(\Gamma\)-like collection closed under finite unions
and \(\mc A\subseteq\mc B\).
Then \(\alpha_1(\mc A,\mc B)\) is equivalent to
\(\alpha_{1.1}(\mc A,\mc B)\).
\end{theorem}

\begin{proof}
Let \(A_n\in\mc A\) and assume \(\alpha_{1.1}(\mc A,\mc B)\).
To apply the assumption, we will define a pairwise disjoint
collection \(\{A_n':n<\omega\}\). First let \(0'=0\) and \(A_0'=A_0\).
Then suppose \(m'\geq m\) and \(A_m'\subseteq A_{m'}\subseteq\bigcup_{i\leq m}A_i'\) 
are defined for all \(m\leq n\).

If \(A_k\setminus\bigcup_{m\leq n}A_m'\) is finite for \(k>n'\), let
\(B=\bigcup_{m\leq n'}A_m\in\mc A\subseteq\mc B\). This \(B\) then witnesses 
\(\alpha_1(\mc A,\mc B)\) since \(A_k\setminus B\) is finite for all \(k<\omega\).

Otherwise pick the minimal \((n+1)'>n\) where 
\(A_{n+1}'=A_{(n+1)'}\setminus\bigcup_{m\leq n}A_m'\) is infinite.
It follows that \(A_{n+1}'\subseteq A_{(n+1)'}\subseteq \bigcup_{m\leq n+1}A_m'\).
By construction, \(\{A_n':n<\omega\}\) is a pairwise disjoint collection of
members of \(\mc A\), and we may
apply \(\alpha_{1.1}(\mc A,\mc B)\) to obtain \(B\in\mc B\) where
\(A_n'\setminus B\) is finite for all \(n<\omega\).

@@

Another fractional version of these \(\alpha\)-principles is given as
\(\alpha_{1.5}\) in [Nyikos 92], defined in general as follows.

\begin{definition}
Let \(\alpha_{1.5}(\mc A,\mc B)\) be the assertion that when \(A_n\in\mc A\)
and \(\{A_n:n<\omega\}\) is pairwise disjoint, then there exists \(B\in\mc B\)
such that \(A_n\cap B\) is cofinite in \(A_n\) for infinitely-many \(n<\omega\).
\end{definition}

It's immediate from their definitions that 
\(\alpha_{1.1}(\mc A,\mc B)\) implies \(\alpha_{1.5}(\mc A,\mc B)\), which
implies \(\alpha_{3.1}(\mc A,\mc B)\).
Nyikos originally showed that \(\alpha_{1.5}(\Gamma_{X,x},\Gamma_{X,x})\) implies
\(\alpha_2(\Gamma_{X,x},\Gamma_{X,x})\); this result generalizes as follows.

\begin{theorem}
Let \(\mc A\) be a \(\Gamma\)-like collection closed under finite unions.
Then \(\alpha_{1.5}(\mc A,\mc B)\) implies
\(\alpha_{2}(\mc A,\mc B)\).
\end{theorem}

\begin{proof}
We assume \(\alpha_{1.5}(\mc A,\mc B)\) and demonstrate
\(\alpha_{2.1}(\mc A,\mc B)\), which is equivalent to
\(\alpha_{2}(\mc A,\mc B)\) by Proposition \ref{pointOne}.
So let \(A_n\in\mc A\) such that \(\{A_n:n<\omega\}\) is pairwise-disjoint.

We may partition each \(A_n\) into \(\{A_{n,m}:m<\omega\}\) with
\(A_{n,m}\in\mc A\) for all \(m<\omega\). 
Let \(A_n'=\bigcup\{A_{i,j}:i+j=n\}\in\mc A\);
since \(\{A_n':n<\omega\}\) is pairwise disjoint, we may apply
\(\alpha_{1.5}(\mc A,\mc B)\) to obtain \(B\in\mc B\) where
\(A_n'\cap B\) is cofinite in \(A_n'\) for infinitely-many \(n<\omega\).

Then for \(n<\omega\), choose \(N\geq n\) with \(A_N'\cap B\) cofinite in \(A_N'\).
Then \(A_{n,N-n}\subseteq A_N'\), 
so \(A_{n,N-n}\cap B\) is cofinite in \(A_{n,N-n}\), in particular,
\(A_{n,N-n}\cap B\) is infinite.
Therefore \(A_n\cap B\) is infinite, and we have shown
\(\alpha_{2.1}(\mc A,\mc B)\).
\end{proof}

\begin{corollary}
Let \(\mc A\) be a \(\Gamma\)-like collection closed under finite unions.
Then \(\alpha_x(\mc A,\mc B)\) implies \(\alpha_y(\mc A,\mc B)\) for
\(1<x\leq y\). Additionally, if \(\mc A\subseteq\mc B\), then 
\(\alpha_x(\mc A,\mc B)\) implies \(\alpha_y(\mc A,\mc B)\) for
\(1\leq x\leq y\).
\end{corollary}

@@

\begin{question}
Let \(\mc A\) be strongly-\(\Gamma\)-like.
Is it consistent that \(\alpha_2(\mc A,\mc A)\)
implies \(\alpha_1(\mc A,\mc A)\)?
\end{question}

@@

The author would like to thank Alan Dow, Jared Holshouser,
and Alexander Osipov for various discussions related to this paper.
